Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2021-06-14 manuscript.tex
Remove weasel words & activate passive sentences [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/9a5eb86d587fe4312a9e8baa931a147ef99ec199.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/9a5eb86d587fe4312a9e8baa931a147ef99ec199/checks
https://github.com/habi/acinar-analysis-manuscript/runs/935404349

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...462370801e58f3852f21ccde9145286c09202383}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/462370801e58f3852f21ccde9145286c09202383/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/462370801e58f3852f21ccde9145286c09202383}{habi/acinar-analysis-manuscript@4623708}
on June 8, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(4623708)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(4623708)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

Pulmonary acini represent the functional gas-exchanging units of the lung.
Due to technical limitations, individual acini cannot be identified on microscopic lung sections.
To overcome these limitations, we imaged the right lower lobes of instillation-fixed rat lungs from postnatal days P4, P10, P21, and P60 at the TOMCAT beamline of the Swiss Light Source synchrotron facility at a voxel size of 1.48 μm.
Individual acini were segmented from the three-dimensional data by closing the airways at the transition from conducting to gas exchanging airways.
For a subset of acini (N=268), we followed the acinar development by stereologically assessing their volume and their number of alveoli.
We found that the mean volume of the acini increases 23 times during the observed time-frame.
The coefficients of variation dropped from 1.26 to 0.49 and the difference between the mean volumes of the fraction of the 20 \% smallest to the 20 \% largest acini decreased from a factor of 27.26 (day 4) to a factor of 4.07 (day 60), i.e.~shows a smaller dispersion at later time points.
The acinar volumes show a very large variation early in lung development and homogenize during maturation of the lung by reducing their size distribution by a factor of 7 until adulthood.
The homogenization of the acinar sizes hints at an optimization of the gas-exchange region in the lungs of adult animals and that acini of different size are not evenly distributed in the lungs.
This likely leads to more homogeneous ventilation at later stages in lung development.

@@

Lung development starts with the formation of the two lung buds.
During the embryonic stage (prenatal days E11--E13 in rats) the major airways and the pleura are formed.
During the pseudoglandular stage (E13--E18.5) most of the remaining airway generations are formed by branching morphogenesis.
Only few additional branches are added during the later stages (canalicular: E18.5--E20) and saccular: E20 to postnatal day P4) {[}\protect\hyperlink{ref-TsNwin2E}{1}{]}.
Epithelial differentiation during the canalicular stage and angiogenetic activation of the capillaries lead to the first functional air-blood barriers in the lung {[}\protect\hyperlink{ref-ODQCVC23}{2}{]}.
The saccular stage represents an intermediate stage where branching morphogenesis ceases and the developmental program switches to the phase of alveolarization.
During alveolarization, 90~\% of the adult gas-exchange area is formed by lifting off new septa from the existing gas-exchange surface {[}\protect\hyperlink{ref-auNiE56G}{3}--\protect\hyperlink{ref-XQLPrPTJ}{5}{]}.
Alveolarization itself is divided into two distinct phases, the so-called classical (P4--P21) and continued alveolarization (P14 to approximately P60) {[}\protect\hyperlink{ref-wnl86DEM}{6},\protect\hyperlink{ref-d1YV8oOK}{7}{]}.
During classical alveolarization, new septa are formed starting from preexisting immature septa which contain a double-layered capillary network.
During continued alveolarization, new septa are formed from preexisting mature septa.
In general, during alveolarization the structure of the alveolar septa is changed to increase the efficiency of the lung {[}\protect\hyperlink{ref-wnl86DEM}{6},\protect\hyperlink{ref-d1YV8oOK}{7}{]}.

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, single or very few acini were delineated and extracted from manually traced serial sections {[}\protect\hyperlink{ref-c3aIB7f9}{17},\protect\hyperlink{ref-CPhMy3Ie}{18}{]} or sectioned silicone casts {[}\protect\hyperlink{ref-1GJECnKBu}{19}{]} with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~{[}\protect\hyperlink{ref-C6OaY84D}{20}{]}.
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Haberthür et al.~{[}\protect\hyperlink{ref-7YLeeyu}{21}{]} analyzed the volume, surface area, and number of alveoli of 43 individual acini of three adult rats (day 60) in tomographic datasets obtained by synchrotron radiation based X-ray tomographic microscopy.
Similarly, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

Due to the lack of precise and complete three-dimensional data, simulations of airflow in the lung are based on relatively simple acinar models.
The question of how well these models represent lung physiology remains open until the necessary data and physiologically correct models are available.
To contribute to the validation of the computational fluid dynamics simulations, we extracted individual acini and determined their volume and their number of alveoli throughout rat lung development.

@@

While the global volume of the acini is well known and studied, only little is known about the range of the volume of individual acini throughout lung development.
We developed the hereby presented method to analyze large amounts of acini over the course of postnatal lung development in a time-efficient manner since no data on large amounts of acini is available up to now.
Our aim was to understand the size distribution of the acini during lung development including the contribution of the rather irregular shape of the alveoli.
The observed range \emph{de}creased during postnatal lung development by a factor of 6--7, which represents a completely unexpected result.
The obtained data are essential for further investigations, like the influence of the size of the acini on ventilation or gas-washout, or the influence of the irregular alveolar shape on partivle deposition inside individual acini, respectively.
E.g., we are currently using our computational fluid dynamics model to simulate exactly this influence (own unpublished results and {[}\protect\hyperlink{ref-eb0gk6VO}{13}{]}).

@@

In the present study, a superset of the animals described by Haberthür et al.~{[}\protect\hyperlink{ref-7YLeeyu}{21}{]} were used.
Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]} described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement) {[}\protect\hyperlink{ref-199ALtdJt}{23}{]} for the ethical use and the reduction of the number of animals sacrificed for science.

Briefly, we extracted the lungs of Sprague-Dawley rats after having deeply anesthetized them with a mixture of medetomidine, midazolam, and fentanyl {[}\protect\hyperlink{ref-IdRLVLjl}{24}{]}.
The rats were euthanized by exsanguination during the removal of the lung.
The lungs were fixed with 2.5~\% glutaraldehyde by intratracheal instillation and kept under constant pressure (20 cm H\textsubscript{2}O) during fixation to prevent recoiling of the tissue.
The samples were prepared for tomographic imaging by post-fixation with 1~\% osmium tetroxide and staining with 4~\% uranyl acetate, as well as by dehydration in a graded series of ethanol and embedding in paraffin.
In total we assessed 12 animals, on postnatal days P4, P10, P21, and P60 (N=3 per day.
From now on, we will omit the postnatal prefix and only mention the day.).

@@

Tomographic imaging was performed at the \href{https://www.psi.ch/sls/tomcat/}{TOMCAT} beamline (A beamline for TOmographic Microscopy and Coherent rAdiology experimenTs) {[}\protect\hyperlink{ref-knMAkPPz}{25},\protect\hyperlink{ref-Cp0VRmEj}{26}{]} of the \href{https://www.psi.ch/sls/}{Swiss Light Source} at the \href{https://www.psi.ch/}{Paul Scherrer Institute} in Villigen, Switzerland.
The samples were scanned at an X-ray energy of 20.0~keV.
After penetrating the sample, the X-rays were converted into visible light by either a 20~μm-thick LuAG:Ce or 18~μm-thick YAG:Ce scintillator screen (both from \href{http://www.crytur.cz/}{Crytur}, Turnov, Czech Republic), depending on the date of experiments.
The resulting visible light was magnified using a 10-times magnifying, diffraction-limited microscope lens and recorded with a 2048 × 2048 pixel CCD camera (\href{https://www.pco.de/specialized-cameras/pco2000/}{pco.2000}, PCO, Kelheim, Germany) with 14 bits dynamic range operated in 2 by 2 binning mode.
As a result, in a single field of view, we were able to image a sample volume of a cube of 1.5~mm side length with a voxel side length of 1.48~μm, with the exposure time of the single projections varying between 160 and 200~ms.

Since our samples were substantially larger than the field of view of the detector, we applied the so-called wide-field scanning method {[}\protect\hyperlink{ref-VELl7OiR}{27}{]} to increase the field of view horizontally.
For each sub-scan, we recorded 3578 projections and laterally stitched them so that their combined field of view covered the whole sample width.
Additionally, two or three such wide-field scans were stacked vertically to match the sample height.

@@

The stereological estimation of the alveolar number was performed according to the standards for quantitative assessment of lung structure from the \href{http://www.thoracic.org/}{American Thoracic Society} {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]} to guarantee accurate and unbiased results.
As a basis for the estimation the volumes of the lung lobes were first determined by water displacement {[}\protect\hyperlink{ref-KGbSQovR}{29}{]}.
The lobe-volume was estimated a second time by the Cavalieri method.
Both volumes were used to calculate the shrinkage, which was used for the correction of the volume data {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}.

To perform the stereological analysis, each DICOM file from the isolated acini was sampled into a systematic random set of single images using a MATLAB script {[}\protect\hyperlink{ref-12Z2YPzm8}{30}{]}.
The stereological assessment was performed with the \href{http://stepanizer.com/}{STEPanizer} {[}\protect\hyperlink{ref-nPoQ2EIB}{31}{]}, an easy-to-use software tool for the stereological assessment of digital images.
Details of the stereological assessment have previously been described {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Briefly, we counted the appearance or disappearance of alveolar wall interruptions.
Under the assumption that these only occur in the region of the alveolar mouth opening and correspond to alveolar entrance rings {[}\protect\hyperlink{ref-QiAxY2i3}{32}{]} we counted said interruptions on paired images spanning the whole acinus volume.
By using the disector method {[}\protect\hyperlink{ref-FJ9FoB4m}{33}{]} we were thus able to estimate the number of alveoli in 268 of the 701 exported acini.
The acini we analyzed were subsampled from the full data by systematic uniform random sampling {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]}.
Briefly, a defined fraction of the acini of one animal were analyzed, e.g.~one third.
During segmentation, the acini were numbered.
Based on these numbers every third acinus was selected, randomly starting with the first, second, or third one.

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle {[}\protect\hyperlink{ref-FE9HLB4f}{34}{]}) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available on request.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{35}{]} notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub {[}\protect\hyperlink{ref-V87xbt0b}{36}{]}.
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amhrein et al.~{[}\protect\hyperlink{ref-o21zxPIu}{37}{]}.

Usually, we performed a Shapiro-Wilk test for normality {[}\protect\hyperlink{ref-IkHrgIj3}{38}{]} to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test {[}\protect\hyperlink{ref-MyzxBJ57}{39}{]}.
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples {[}\protect\hyperlink{ref-ON1Bppkk}{40}{]} was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python {[}\protect\hyperlink{ref-8Miti2Gz}{41}{]} or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences (Šidák-corrected p-value smaller than 0.00167 {[}\protect\hyperlink{ref-GtCynqsI}{42}{]}) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are smaller than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

The counts per volume, which correspond to the number of alveoli per acinus volume are one of several measure for influencing the complexity of the single acini.
The number of alveoli per acinus show highly significant differences between all possible combinations of days except between days 10 and 21 (p=0.7).
All other days have a p-value smaller than 5e-5, this being the p-value between days 4 and 21.
The statistical analysis is equal to what is briefly described for the entrance ring count above.
The counts per volume for all animals except 04A are non-normally distributed, but all are independent (p=2.9e-21).

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10}{]} and that information drawn from one lobe is representative of the whole lung {[}\protect\hyperlink{ref-I9TmP6IU}{43}{]}.
Furthermore, based on stereological data, Zeltner et al.~{[}\protect\hyperlink{ref-I9TmP6IU}{43}{]} were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy {[}\protect\hyperlink{ref-Zv1XcgAO}{44}{]}.
A first analysis of regional difference shows similar results as mentioned above.

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...9a5eb86d587fe4312a9e8baa931a147ef99ec199}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/9a5eb86d587fe4312a9e8baa931a147ef99ec199/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/9a5eb86d587fe4312a9e8baa931a147ef99ec199}{habi/acinar-analysis-manuscript@9a5eb86}
on June 14, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(9a5eb86)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(9a5eb86)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

Pulmonary acini represent the functional gas-exchanging units of the lung.
Due to technical limitations, individual acini cannot be identified on microscopic lung sections.
To overcome these limitations, we imaged the right lower lobes of instillation-fixed rat lungs from postnatal days P4, P10, P21, and P60 at the TOMCAT beamline of the Swiss Light Source synchrotron facility at a voxel size of 1.48 μm.
Individual acini were segmented from the three-dimensional data by closing the airways at the transition from conducting to gas exchanging airways.
For a subset of acini (N=268), we followed the acinar development by stereologically assessing their volume and their number of alveoli.
We found that the mean volume of the acini increases 23 times during the observed time-frame.
The coefficients of variation dropped from 1.26 to 0.49 and the difference between the mean volumes of the fraction of the 20 \% smallest to the 20 \% largest acini decreased from a factor of 27.26 (day 4) to a factor of 4.07 (day 60), i.e.~shows a smaller dispersion at later time points.
The acinar volumes show a large variation early in lung development and homogenize during maturation of the lung by reducing their size distribution by a factor of 7 until adulthood.
The homogenization of the acinar sizes hints at an optimization of the gas-exchange region in the lungs of adult animals and that acini of different size are not evenly distributed in the lungs.
This likely leads to more homogeneous ventilation at later stages in lung development.

@@

Lung development starts with the formation of the two lung buds.
During the embryonic stage (prenatal days E11--E13 in rats) the major airways and the pleura are formed.
During the pseudoglandular stage (E13--E18.5) most of the remaining airway generations are formed by branching morphogenesis.
Only a small amount of additional branches are added during the later stages (canalicular: E18.5--E20) and saccular: E20 to postnatal day P4) {[}\protect\hyperlink{ref-TsNwin2E}{1}{]}.
Epithelial differentiation during the canalicular stage and angiogenetic activation of the capillaries lead to the first functional air-blood barriers in the lung {[}\protect\hyperlink{ref-ODQCVC23}{2}{]}.
The saccular stage represents an intermediate stage where branching morphogenesis ceases and the developmental program switches to the phase of alveolarization.
During alveolarization, 90~\% of the adult gas-exchange area is formed by lifting off new septa from the existing gas-exchange surface {[}\protect\hyperlink{ref-auNiE56G}{3}--\protect\hyperlink{ref-XQLPrPTJ}{5}{]}.
Alveolarization itself is divided into two distinct phases, the so-called classical (P4--P21) and continued alveolarization (P14 to approximately P60) {[}\protect\hyperlink{ref-wnl86DEM}{6},\protect\hyperlink{ref-d1YV8oOK}{7}{]}.
During classical alveolarization, new septa are formed starting from preexisting immature septa which contain a double-layered capillary network.
During continued alveolarization, new septa are formed from preexisting mature septa.
In general, during alveolarization the structure of the alveolar septa is changed to increase the efficiency of the lung {[}\protect\hyperlink{ref-wnl86DEM}{6},\protect\hyperlink{ref-d1YV8oOK}{7}{]}.

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, only a small amount of acini were delineated and extracted from manually traced serial sections {[}\protect\hyperlink{ref-c3aIB7f9}{17},\protect\hyperlink{ref-CPhMy3Ie}{18}{]} or sectioned silicone casts {[}\protect\hyperlink{ref-1GJECnKBu}{19}{]} with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~{[}\protect\hyperlink{ref-C6OaY84D}{20}{]}.
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Haberthür et al.~{[}\protect\hyperlink{ref-7YLeeyu}{21}{]} analyzed the volume, surface area, and number of alveoli of 43 individual acini of three adult rats (day 60) in tomographic datasets obtained by synchrotron radiation based X-ray tomographic microscopy.
Similarly, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

Due to the lack of precise and complete three-dimensional data, simulations of airflow in the lung are based on simple idealized acinar models {[}\protect\hyperlink{ref-NPtttMC0}{23},\protect\hyperlink{ref-6zXNR8r2}{24}{]}.
The question of how well these models represent lung physiology remains open until the necessary data and physiologically correct models are available.
To contribute to the validation of the computational fluid dynamics simulations, we extracted individual acini and determined their volume and their number of alveoli throughout rat lung development.

@@

While the global volume of the acini is well known and studied, only little is known about the range of the volume of individual acini throughout lung development.
We developed the hereby presented method to analyze large amounts of acini over the course of postnatal lung development in a time-efficient manner since no data on large amounts of acini is available up to now.
Our aim was to understand the size distribution of the acini during lung development including the contribution of the rather irregular shape of the alveoli.
The observed range \emph{de}creased during postnatal lung development by a factor of 6--7, which represents an unexpected result.
The obtained data are essential for further investigations, like the influence of the size of the acini on ventilation or gas-washout, or the influence of the irregular alveolar shape on partivle deposition inside individual acini, respectively.
E.g., we are currently using our computational fluid dynamics model to simulate exactly this influence (own unpublished results and {[}\protect\hyperlink{ref-eb0gk6VO}{13}{]}).

@@

In the present study, a superset of the animals described by Haberthür et al.~{[}\protect\hyperlink{ref-7YLeeyu}{21}{]} were used.
Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]} described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement) {[}\protect\hyperlink{ref-199ALtdJt}{25}{]} for the ethical use and the reduction of the number of animals sacrificed for science.

Briefly, we extracted the lungs of Sprague-Dawley rats after having deeply anesthetized them with a mixture of medetomidine, midazolam, and fentanyl {[}\protect\hyperlink{ref-IdRLVLjl}{26}{]}.
The rats were euthanized by exsanguination during the removal of the lung.
The lungs were fixed with 2.5~\% glutaraldehyde by intratracheal instillation and kept under constant pressure (20 cm H\textsubscript{2}O) during fixation to prevent recoiling of the tissue.
The samples were prepared for tomographic imaging by post-fixation with 1~\% osmium tetroxide and staining with 4~\% uranyl acetate, as well as by dehydration in a graded series of ethanol and embedding in paraffin.
In total we assessed 12 animals, on postnatal days P4, P10, P21, and P60 (N=3 per day.
From now on, we will omit the postnatal prefix and only mention the day.).

@@

Tomographic imaging was performed at the \href{https://www.psi.ch/sls/tomcat/}{TOMCAT} beamline (A beamline for TOmographic Microscopy and Coherent rAdiology experimenTs) {[}\protect\hyperlink{ref-knMAkPPz}{27},\protect\hyperlink{ref-Cp0VRmEj}{28}{]} of the \href{https://www.psi.ch/sls/}{Swiss Light Source} at the \href{https://www.psi.ch/}{Paul Scherrer Institute} in Villigen, Switzerland.
The samples were scanned at an X-ray energy of 20.0~keV.
After penetrating the sample, the X-rays were converted into visible light by either a 20~μm-thick LuAG:Ce or 18~μm-thick YAG:Ce scintillator screen (both from \href{http://www.crytur.cz/}{Crytur}, Turnov, Czech Republic), depending on the date of experiments.
The resulting visible light was magnified using a 10-times magnifying, diffraction-limited microscope lens and recorded with a 2048 × 2048 pixel CCD camera (\href{https://www.pco.de/specialized-cameras/pco2000/}{pco.2000}, PCO, Kelheim, Germany) with 14 bits dynamic range operated in 2 by 2 binning mode.
As a result, in a single field of view, we were able to image a sample volume of a cube of 1.5~mm side length with a voxel side length of 1.48~μm, with the exposure time of the single projections varying between 160 and 200~ms.

Since our samples were larger than the field of view of the detector, we applied the so-called wide-field scanning method {[}\protect\hyperlink{ref-VELl7OiR}{29}{]} to increase the field of view horizontally.
For each sub-scan, we recorded 3578 projections and laterally stitched them so that their combined field of view covered the whole sample width.
Additionally, two or three such wide-field scans were stacked vertically to match the sample height.

@@

The stereological estimation of the alveolar number was performed according to the standards for quantitative assessment of lung structure from the \href{http://www.thoracic.org/}{American Thoracic Society} {[}\protect\hyperlink{ref-dNc8FfNn}{30}{]} to guarantee accurate and unbiased results.
As a basis for the estimation the volumes of the lung lobes were first determined by water displacement {[}\protect\hyperlink{ref-KGbSQovR}{31}{]}.
The lobe-volume was estimated a second time by the Cavalieri method.
Both volumes were used to calculate the shrinkage, which was used for the correction of the volume data {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}.

To perform the stereological analysis, each DICOM file from the isolated acini was sampled into a systematic random set of single images using a MATLAB script {[}\protect\hyperlink{ref-12Z2YPzm8}{32}{]}.
The stereological assessment was performed with the \href{http://stepanizer.com/}{STEPanizer} {[}\protect\hyperlink{ref-nPoQ2EIB}{33}{]}, an easy-to-use software tool for the stereological assessment of digital images.
Details of the stereological assessment have previously been described {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Briefly, we counted the appearance or disappearance of alveolar wall interruptions.
Under the assumption that these only occur in the region of the alveolar mouth opening and correspond to alveolar entrance rings {[}\protect\hyperlink{ref-QiAxY2i3}{34}{]} we counted said interruptions on paired images spanning the whole acinus volume.
By using the disector method {[}\protect\hyperlink{ref-FJ9FoB4m}{35}{]} we were thus able to estimate the number of alveoli in 268 of the 701 exported acini.
The acini we analyzed were subsampled from the full data by systematic uniform random sampling {[}\protect\hyperlink{ref-dNc8FfNn}{30}{]}.
A defined fraction (i.e.~one third) of all the extracted acini of one animal were analyzed.
During segmentation, the acini were numbered.
Based on these numbers every third acinus was selected, randomly starting with the first, second, or third one.

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle {[}\protect\hyperlink{ref-FE9HLB4f}{36}{]}) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets {[}\protect\hyperlink{ref-7YLeeyu}{21}{]}.
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available on request.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} {[}\protect\hyperlink{ref-pQ6Wbz73}{37}{]} notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub {[}\protect\hyperlink{ref-V87xbt0b}{38}{]}.
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amhrein et al.~{[}\protect\hyperlink{ref-o21zxPIu}{39}{]}.

Usually, we performed a Shapiro-Wilk test for normality {[}\protect\hyperlink{ref-IkHrgIj3}{40}{]} to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test {[}\protect\hyperlink{ref-MyzxBJ57}{41}{]}.
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples {[}\protect\hyperlink{ref-ON1Bppkk}{42}{]} was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python {[}\protect\hyperlink{ref-8Miti2Gz}{43}{]} or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences (Šidák-corrected p-value smaller than 0.00167 {[}\protect\hyperlink{ref-GtCynqsI}{44}{]}) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are smaller than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

The counts per volume---which correspond to the number of alveoli per acinus volume---are one measures for the complexity of the single acini.
The number of alveoli per acinus show highly significant differences between all possible combinations of days except between days 10 and 21 (p=0.7).
All other days have a p-value smaller than 5e-5, this being the p-value between days 4 and 21.
The statistical analysis is equal to what is briefly described for the entrance ring count above.
The counts per volume for all animals except 04A are non-normally distributed, but all are independent (p=2.9e-21).

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10}{]} and that information drawn from one lobe is representative of the whole lung {[}\protect\hyperlink{ref-I9TmP6IU}{45}{]}.
Furthermore, based on stereological data, Zeltner et al.~{[}\protect\hyperlink{ref-I9TmP6IU}{45}{]} were able to show that the lung parenchyma is rather homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~{[}\protect\hyperlink{ref-RGBeCf8v}{22}{]} observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, for example between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy {[}\protect\hyperlink{ref-Zv1XcgAO}{46}{]}.
A first analysis of regional difference shows similar results as mentioned above.
